On The Difficulty of Computing Non-Commutative Signature Gröbner Bases

### 2022-02-14 article_issac.tex
début preuve proposition context free

\begin{proposition}
  Leading monomials of syzygies of $F=\{g_1,\dots,g_{18}\}$ do not form an context free language.
  \label{prop:exp_pas_regulier}
\end{proposition}
\begin{proof}
  For every $n \geq 1$, $g_{11}a^nb^ne$ is the leading monomial of the syzygy:

>>>>

\begin{proposition}
  Leading monomials of syzygies of $F=\{g_1,\dots,g_{18}\}$ do not form an context free language.
  \label{prop:exp_pas_regulier}
\end{proposition}
\begin{proof}
  For every $n \geq 1$, $g_{18}a^nb^nc^ne$ is the leading monomial of the syzygy following from the adaptation
  of the previous sequence of equalities for higher numbers of $a$'s, $b$'s and $c$'s.

  Now, let us show that leading monomials of syzygies of the form $\varepsilon_{18}a^nb^mc^ke$ are such that $n = m=k$.
  Let $\sigma$ be such a syzygy and let $G$ be the graph whose vertices are elements of $\left\langle X \right\rangle$, and there is an edge between vertices of the form $l\omega r$ and $l\omega'r$ as soon as $lg_ir$ occurs with a nonzero coefficient in $\sigma$ and $g_i=\omega - \omega'$. Let $W$ be the set of vertices of the following form:
  \begin{align*}
    & ssa^nb^ne, \\
    & tc^pa^qxa^rd^nb^me, \\
    & tc^pa^qd^rxd^nb^me, \\
    & tc^pa^qya^rd^nb^me, \\
    & tc^pa^qd^ryd^nb^me, \\
    & tc^pza^qb^re, \\
    & tc^pd^qzb^re
  \end{align*}
  Each vertex in $W$ is related to at most two vertices, which also belong to $W$. When an undirected graph is such that each vertex is related to at most two vertices, each connected component is either a chain or cycle, so that the connected component of $ssa^nb^me$ in $G$ is either a chain or a cycle.
  But since $\sigma$ is a syzygy, this connected component has to be a cycle, which implies $n = m$.
  Let $\mathscr{C}$ be defined by:
$$\mathscr{C}=\{l\varepsilon_ir \in \MM(\mathscr{F}_r) \;|\; l \LM(g_i) r \mbox{ does not contain } \infty\}.$$
   Let $\mathscr{A}$ be the set of leading monomials of syzygies of $F$.
   Now, we can conclude that the set $\mathscr{A} \cap \mathscr{C}$ does not form a regular language.
  Indeed, by way of contradiction, assume that $\mathscr{A}\cap \mathscr{C}$ is a regular language. Then its
  intersection with the regular language $\{g_{11}a^nb^me \;|\; n, m \geq 1\}$ also forms regular language.
  We reach a contradiction since we proved that this intersection is $\{g_{11}a^nb^ne \;|\; n \geq 1 \}$, which does not form a regular language.
\end{proof}

\begin{cor}
There is a set $F'$ of polynomials which
admits a finite Gröbner basis
and whose set of leading monomials of syzygies does not form a context free language. 
\end{cor}
\begin{proof}
  We proceed as in the previous section: a counter example is given by the ideal
  generated by $g_1,\dots,g_{18}$ and polynomials of the form $\infty-d$, for letters
  $d$ different from $\infty$.
\end{proof}
